### 2025-08-04 manuscript_res.tex
add some score plots

\section{Introduction}

Every year the seasonal influenza outbreak burdens the public health system by 
infecting millions, causing an influx of primary care visits and 
hospitalizations and leading to between 290,000 and 650,000 deaths worldwide 
\cite[]{whoflufact2023}. Molinari et al. \cite[]{molinari2007annual} estimated 
the United States' annual economic burden from medical costs, loss of income, 
and deaths to be over \$87 billion. Accurate forecasting of infectious diseases 
can inform public decision making and ease the burden of an outbreak 
\cite[]{turtle2021accurate, lutz2019applying}.
There is a growing consensus that disease forecasts should be probabilistic in 
nature \cite[]{gneiting2014probabilistic, bracher2021evaluating}, and it has 
been shown that reporting forecast uncertainty along with predictions may lead 
to better decision making 
\cite[]{ramos2013probabilistic, joslyn2012uncertainty, winkler1971probabilistic}.

% \jarad{Test}
% \spencer{Test}

To better inform public decision making regarding the flu epidemic, in 2013 the 
United States Centers for Disease Control and Prevention (CDC) organized a 
national flu forecasting competition, also known as FluSight 
\cite[]{biggerstaff2016results, mathis2024evaluation,cdcfluforecasting2024}. 
Originally, over a dozen teams of researchers from academic and industry 
backgrounds participated in FluSight by contributing their own forecast models. 
Besides the 2020 season --or the flu season spanning the fall of 2020 and the 
winter of 2021-- FluSight has been operated annually and researchers outside 
the CDC have been invited to participate. Initially the target data for 
forecasts was influenza-like illness (ILI) data. ILI is the proportion of 
patients who meet a healthcare provider and who display flu like symptoms, 
and ILI data has been available at the state and national level since the 2010 
flu season \cite[]{cdcfluview2023,cdc2024fluviewportal}. 
The collaborative ILI forecasting effort has led to a number of modeling 
developments in flu forecasting 
\cite[see references therein for more examples]{mcandrew2021adaptively, 
osthus2021multiscale, osthus2019dynamic, ulloa2019}, and in their paper's 
introduction, Osthus et al.  \cite[]{osthus2019dynamic} categorized the most 
commonly used flu forecasting models into four classes including mechanistic 
models based on differential equation compartmental models, agent based models 
based on population simulation, machine learning/regression models including 
data driven machine learning and statistical models, and data assimilation 
models which are constructed by assimilating mechanistic models into a 
probabilistic framework. An additional forecast model used in FluSight 
involves the combination of several forecasts into a single ensemble forecast, 
which has been shown to perform well relative to individual models 
\cite[]{mcandrew2021adaptively, ray2020ensemble, yamana2016superensemble}.

The administration of FluSight saw few changes during the first seven seasons
between 2013 and 2019, 
but the onset of the COVID-19 pandemic and subsequent developments for COVID-19 
forecasting led to major modifications. As a result of the COVID-19 pandemic 
which began during the 2019 flu season, the typical flu outbreak behavior was 
altered between the 2019 and 2022 seasons \cite[]{mathis2024evaluation}. 
The COVID-19 pandemic led to the creation of the Health and Human Services 
(HHS) Patient Impact and Hospital Capacity Data System 
\cite[]{healthdata2024covidts} which contains COVID-19 and flu hospitalization 
data, and the COVID-19 Forecast Hub was founded. The COVID-19 Forecast Hub 
was based on FluSight but with certain major adjustments including how the 
forecast uncertainty is represented and the addition of the weekly publication 
of a multi-model ensemble forecast as the official forecast of the CDC 
\cite[]{bracher2021evaluating, Cramer2022-hub-dataset}. Using estimated 
quantiles for representing forecast uncertainty and creating a multi-model 
ensemble are both aspects of the COVID-19 Forecast Hub which were adopted by 
the flu forecast competition. Additionally the target of the flu forecasts 
changed from being ILI data to being HHS hospitalization data, which reports 
the number of hospitalizations due to a laboratory confirmed flu infection  
\cite[]{mathis2024evaluation,healthdata2024covidts}. This is as a result of 
having COVID-19 cases in the population making ILI data, already only a proxy 
for flu behavior, more difficult to interpret.

The contribution of this manuscript is to introduce a two component framework 
for modeling HHS hospitalization forecasts where hospitalization data and 
years of ILI data are used to inform forecast models. The first modeling 
component is a model of ILI data and the second is a model of hospitalization 
data with ILI as a predictive covariate.
Herein we use ILI models similar to those in Osthus et al. 
\cite[]{osthus2019dynamic} and Ulloa \cite[]{ulloa2019} for ILI forecasting.
The model of Osthus et al. \cite[]{osthus2019dynamic} is a combined data 
assimilation and statistical regression model which involves a compartmental 
model in a probabilistic framework. Their model also includes an additional 
component for capturing a systematic discrepancy between the deterministic 
part of the model and the actual data, an idea which was first introduced by 
Kennedy and O'Hagan \cite[]{kennedy2001bayesian}. The model in Ulloa 
\cite[]{ulloa2019} is a Bayesian hierarchical regression model with an 
underlying function intended to capture the trajectory of the seasonal 
ILI data. Herein, we provide a framework under which discrepancy modeling may 
be used along with a general function modeling ILI data, and we show the 
effectiveness of including discrepancy modeling during certain periods of 
the flu season.

In line with the newer FluSight standard of forecasting hospitalizations, we 
model hospitalizations as a linear function of ILI. Thus forecasts produced 
herein target flu hospitalizations and are a mapping of ILI forecasts to 
hospitalizations. This allows for ILI data from many seasons to be exploited 
and for ILI forecasts to assist in forecasting hospitalizations, which has 
fewer seasons of data than ILI. Several modeling schemes and their forecasts 
for the 2023 flu hospitalization season are compared, and it is shown that the 
modeling decisions produce good forecast results for different states or times 
during the flu season.

In section \ref{sec:data} we review the ILI and hospitalization data provided 
by the CDC and targeted by FluSight. 
In section \ref{sec:functions} the modeling framework contributed by this 
manuscript is given. In the same section, functions similar to those used by 
Osthus et al. \cite[]{osthus2019dynamic} and Ulloa \cite[]{ulloa2019} are 
defined.
 These functions are the susceptible-infectious-recovered (SIR) compartmental 
 model and the asymmetric Gaussian (ASG) function  respectively. Model fitting 
 and implementation are described at the end of the section. Section 
 \ref{sec:simulation2} includes a simulation study where four ILI forecast 
 models and their use in forecasting hospitalizations are compared. Commonly 
 used proper scoring rules \cite[]{gneiting2007strictly}, which are also 
 introduced and defined in section \ref{sec:simulation2}, are used for 
 comparing the forecasts. Forecasting of the 2023 flu outbreak along with 
 assessment and comparison under several selected models is performed in 
 section \ref{sec:analysis}. 
 Finally, the manuscript is concluded in section \ref{sec:conclusion} with 
 general observations and some discussion.

 \section{Flu outbreak data} \label{sec:data}
In this section we introduce and define ILI and hospitalization data and 
visually evaluate the data. ILI and hospitalization data have been the object 
of forecasting for FluSight with ILI being the target for the first seven 
seasons and hospitalizations being the target since the 2022 season. Both of 
these data were collected at the state, territorial, and national level and 
were reported at least weekly. Overall the data is reported for 53 locations 
including the 50 US states, the District of Columbia (DC), Puerto Rico (PR), 
and at the US national level. We will refer to forecast targets throughout this 
manuscript. A target is the specific horizon, 1, 2, 3, or 4-weeks ahead, for a 
specific location and week during the season.

\subsection{Influenza-like illness data}
The US Outpatient Influenza-like Illness Surveillance Network (ILINet) collects 
information on respiratory illness from outpatient visits to health care 
providers. Over 3,400 outpatient health care providers in all 50 US states, PR, 
DC, and the US Virgin Islands report each week the total number of outpatient 
visits along with the number of ILI cases. An ILI case is defined as a 
``fever (temperature of $100^{\circ}$F[$37.8^{\circ}$C] or greater) and a cough 
and/or a sore throat." Prior to the 2021 season, the definition included 
``without a known cause other than influenza" \cite[]{cdcfluview2023}. Because 
other illnesses such as COVID-19, RSV, and the common cold may induce similar 
respiratory symptoms, ILI may include patients infected with some disease 
other than influenza. To know whether or not a sick patient is infected with 
influenza requires a laboratory test.

In 2013, when FluSight began, the ILI data was the object of the forecasts. 
The data was released publicly at HHS region levels, and forecast teams were 
asked to provide forecasts of several ILI targets on the regional levels 
including season onset, 1-4 week ahead ILI levels, and the week of peak ILI 
activity \cite[]{biggerstaff2016results,mcgowan2019collaborative}. Currently, 
the ILI data is collected by the CDC and published on an online portal for 
viewing at the national, HHS region, census, and state levels 
\cite[]{cdc2024fluviewportal}. To obtain ILI data, we used the \texttt{R} 
package \texttt{cdcfluview} which provides functions for downloading the data 
\cite[]{rudis2021cdcfluview}. Weekly ILI data from the national, HHS region, 
and census levels are available from the 1997 flu season until the current 
season. At the state level, data is available from the 2010 flu season to the 
current season.

The top of figure \ref{fig:us_ili} shows the ILI data at the national level for 
flu seasons 2010 to 2023. For most seasons there are 52 weeks listed, but for 
the 2010, 2015, and 2021 seasons there are 53 weeks because there were 53 
Sundays during those seasons. 
To better align with the flu behavior, week 1 is set as the first week of 
August and week 52 or 53 is the last week in July of the following year. For 
example week 1 of the 2013 season corresponds to the first week of August 2013, 
and week 52 of the same season corresponds to the last week of July 2014. This 
convention is used for the remainder of this manuscript.

\begin{figure}[hbt!]
    \centering
    \includegraphics[scale=.5]{Images/us_ili_seasons.png}

    \centering
    \includegraphics[scale=.4]{Images/us_hospitalizations.png}
    \caption{Percentage of outpatient visits with an influenza-like illness 
    (ILI) in the US for seasons 2010 to 2023 (top) and weekly flu confirmed 
    hospitalization counts at the national level for 2022 and 2023 flu seasons 
    (bottom). Week 1 is the first week of August of the year the flu season 
    begins and the last week of the season is the last week of July of the 
    following year.}
    \label{fig:us_ili}
\end{figure}

Notable from the plots in the top of figure \ref{fig:us_ili} is the regular 
trajectory of the ILI. With the exception of season 2020, the ILI begins low 
at week 1 and increases as the fall and winter progress until the ILI reaches 
a peak. As spring progresses to summer, the ILI decreases to low values. As 
Osthus et al. \cite[]{osthus2019dynamic} point out, there is nearly always 
either a global or local peak at week 22 which typically corresponds to the 
week between Christmas and New Year's day. 
Whether local or global, the ILI holiday peak is generally expected and thought 
to be due to widespread holiday travel, school closure, or other unique social 
behavior   
\cite[]{ewing2017contact, garza2013effect}.
The only seasons when there was not a peak at week 22 were season 2022 where 
the season peak occurred particularly early and season 2020 which was greatly 
influenced by the COVID-19 pandemic.

\subsection{Hospitalization Data}

Hospital admission data, used as the object of FluSight forecasting for the 
2022 and 2023 seasons, is based on the CDC's National Healthcare Safety 
Network (NHSN) dataset entitled \textit{HealthData.gov COVID-19 Reported 
Patient Impact and Hospital Capacity by State Timeseries}. Several targets of 
respiratory illnesses including COVID-19, RSV, and influenza are reported 
weekly by most hospitals in the US. In February 2022 it became mandatory for 
all hospitals to report the number of COVID-19 and influenza hospitalizations, 
and since then reporting of hospitalizations has become widespread. These data 
were updated every Wednesday and Friday according to NHSN guidelines 
\cite[]{healthdata2024covidts}.

The bottom of figure \ref{fig:us_ili} shows the weekly national 
hospitalizations for the 2022 and 2023 flu seasons. These plots show 
similarities to the ILI plots in the top of the figure in that at the early 
weeks of the season hospitalizations are low, but they increase in the fall 
to a peak after which they decrease until the flu outbreak ends. For both the 
2022 and 2023 seasons, the hospitalizations peaked during the same week as 
ILI, and in 2023 that peak occurred during the holiday week 22. Additional 
plots of ILI data and hospitalization
data for select states are included in the supplementary material 
\cite[]{wadsworth2024bas}. These plots show flu outbreak trajectories similar 
to 
those in figure \ref{fig:us_ili} with activity increasing
to a peak and then decreasing.

% \begin{figure}
%     \centering
%     \includegraphics[scale=.5]{Images/us_hospitalizations.png}
%     \caption{Weekly flu confirmed hospitalization counts at the national level for 2022 (left) and 2023 (right) flu seasons}
%     \label{fig:us_hosp}
% \end{figure}

\section{ILI and hospitalization forecast modeling} \label{sec:functions}

% \jarad{Why not start with the ILI model from Osthus et al? It makes a lot more sense to me to start with what has already been done and then expand upon what they have done. Later you have a section called ILI Model, but isn't this the ILI model?}

The typical behavior of the ILI data which starts at lower values in the late 
summer and fall but which increases to a peak, usually in December or January, 
followed by a decline motivates the use of a nonlinear function which follows 
a similar trajectory. Compartmental models are standard mathematical models 
used for modelling disease outbreaks. One important compartmental model is the 
susceptible-infectious-recovered (SIR) model, which is used by some to model 
ILI data \cite[]{osthus2019dynamic, allen2017primer}. Ulloa \cite[]{ulloa2019} 
chose to use the asymmetric Gaussian (ASG) function to model ILI data. The 
SIR and ASG models may both be appropriate to describe ILI behavior over the 
course of a flu season, however there may also be systematic behavior not 
captured by either, necessitating an additional model component to capture 
the discrepancy. In the first part of this section, we present an ILI model 
similar to the model in Osthus et al. \cite[]{osthus2019dynamic}. With some 
generalization, the model may incorporate any appropriate nonlinear function, 
though the focus here is on the SIR and ASG functions which are defined.

With the aim of forecasting hospitalizations, we also introduce a linear 
model of hospitalizations data with ILI as a predictive covariate. To forecast 
hospitalizations, ILI data is first forecasted and the forecast is then 
plugged in as a covariate in the hospitalization model, thus producing 
hospitalization forecasts. The hospitalization model is also defined in this 
section, and the section is concluded with descriptions of selected prior 
distributions, model implementation, and posterior sampling.

\subsection{ILI Model} \label{sec:ili_model}

The proposed model for ILI for any location is given in (\ref{eq:ili_model}). 
Here $ILI_{s,w}$ is the ILI for flu season $s$ and week $w = 1, 2, ..., W$, 
where $W = 52$ or $W = 53$, depending on how many Sundays there are in a given 
season. The ILI is a proportion, so the Beta random variable is a natural 
selection for modeling. Under the parameterization in of the Beta distribution 
used in (\ref{eq:ili_model}) the expected value is $\pi_{s,w}$ and the 
variance is $\pi_{s,w}(1 - \pi_{s,w})/(1 + \kappa_s)$, making $\kappa_s$ a 
scale parameter. The nonlinear function $f_{\theta_s}(w)$ captures the 
trajectory of the ILI, and $\gamma_w$ is a discrepancy term for capturing the 
systematic patterns which $f_{\theta_s}(w)$ does not capture. 
\begin{equation}
\begin{aligned}
    \label{eq:ili_model}
    % \begin{flalign*}
        ILI_{s,w} &\overset{ind}{\sim} \text{Beta}(\pi_{s,w}\kappa_s,\; \kappa_s(1 - \pi_{s,w})) \\
        \text{logit}(\pi_{s,w}) &= f_{\theta_s}(w) + \gamma_w + \upsilon_{s,w}
    % \end{flalign*}
\end{aligned}
\end{equation}
In Osthus et al. \cite[]{osthus2019dynamic} 
$f_{\theta_s}(w) = \text{logit}(I_{s,w})$ where $I_{s,w}$ is the infectious 
compartment of the SIR model from (\ref{eq:sir_diff}) in section 
\ref{sec:sir_func}. In Ulloa \cite[]{ulloa2019} 
$f_{\theta_s}(w) = ASG_{\theta}(w)$ from (\ref{eq:asg_function}) in section 
\ref{sec:asg_func}. In Ulloa \cite[]{ulloa2019} modeling is done hierarchically 
over seasons.

\subsection{Susceptible-Infectious-Recovered (SIR) compartmental model} \label{sec:sir_func}

The SIR compartmental model is a mathematical model used for modeling disease 
outbreaks and was introduced by Kermack and McKendrick 
\cite[]{kermack1927contribution}. Since then, compartmental models have 
become standard for modeling infectious diseases 
\cite[]{allen2008mathematical}, and many extensions have been made and 
studied \cite[for example]{simon2020sir, allen2017primer, van2008deterministic}. 
The SIR mathematical model includes three compartments and assumes that at 
any time $t>0$ every individual in a closed population belongs to exactly 
one compartment. The three compartments are susceptible ($S$), infectious ($I$), 
and recovered ($R$), and their interaction over the course of an outbreak is 
described by the differential equations in (\ref{eq:sir_diff}).

\begin{equation}
    \label{eq:sir_diff}
    \frac{dS}{dt} = -\beta SI, \quad \frac{dI}{dt} = \beta S I - \delta I, \quad \frac{dR}{dt} = \delta I
\end{equation}
Here $S$, $I$, and $R$ represent the proportion of the population in each 
compartment such that $S + I + R = 1$ for all $t$. The trajectory is 
determined by the disease transmission rate $\beta > 0$ and the recovery 
rate $\delta > 0$. Respectively, these may be thought of as the expected 
proportion of susceptible individuals who will be infected by an infectious 
individual and the expected rate of recovery to an immune state for a newly 
infected person. Whether or not a disease outbreak is classified as an 
epidemic is determined by the initial susceptible population $S_0$, or the 
susceptible population at time $0$, and the parameter $\rho = \delta/
\beta$. If $S_0/\rho > 1$, the outbreak is considered an epidemic. It is 
non-epidemic if $S_0/\rho \leq 1$ \cite[]{osthus2019dynamic}. Figure 
\ref{fig:sir_traj} shows the trajectory of the three compartments of an SIR 
model where $S_0$ and $\rho$ were selected to match an outbreak that is an 
epidemic. In the case where $S_0 \leq \rho$, the trajectory for the $I$ 
compartment would never be increasing. The increase to a peak and subsequent 
decrease in the $I$ compartment of figure \ref{fig:sir_traj} suggest it is 
reasonable to model ILI by this compartment. Thus in modeling the ILI data, 
we consider the data to be analogous to the $I$ proportion of the population.

\begin{figure}[hbt!]
    \centering
    \includegraphics[scale=.6]{Images/sir_traj.png}
    \caption{Susceptible-infectious-recovered (SIR) model separated by 
    compartments. The three compartments are the susceptible compartment 
    (left), infectious (center), and recovered (right). In this 
    example, $S_0/\rho > 1$.}
    \label{fig:sir_traj}
\end{figure}

\subsection{Asymmetric Gaussian (ASG) function} \label{sec:asg_func}
The ASG function is another example of a nonlinear function which can 
approximate the trajectory of the flu outbreak. The ASG was previously used 
by Ulloa to model and forecast ILI \cite[]{ulloa2019}, and it has been used 
to model vegetation growth and satellite sensor data 
\cite[]{lewis2020extracting, jonsson2002seasonality, hird2009noise, 
beck2006improved, atkinson2012inter}. The ASG is a modification of the 
asymmetric Gaussian distribution \cite[]{wallis2014two} and is characterized 
by its rise to a peak and fall from that peak which may not occur at the same 
rate, as shown in figure \ref{fig:asg_function}. The ASG function is denoted 
as $ASG_\theta(w)$ where 
$\theta = (\lambda, \nu, \mu, \sigma_1^2, \sigma_2^2)$, $\nu > 0$, 
$\lambda > 0$, $\mu \in (-\infty, \infty)$, $\sigma_1, \sigma_2 > 0$ and 
$w \in (1, ..., W)$ is week. The function is defined in 
(\ref{eq:asg_function}).

\begin{figure}[hbt!]
    \centering
    \includegraphics[scale=.45]{Images/asg_function.png}
    \caption{Example plot of asymmetric Gaussian (ASG) function showing the 
    shape of the function in relation to the parameters 
    $\lambda, \; \eta, \; \mu, \; \sigma_1, \; \text{and} \; \sigma_2$}
    \label{fig:asg_function}
\end{figure}

\begin{equation}
    \label{eq:asg_function}
    ASG_{\theta}(w) = 
    \begin{cases}
        \lambda + (\nu - \lambda) \text{exp}[-(w - \mu)^2/2\sigma^2_1], \;\;\; w < \mu \\
        \lambda + (\nu - \lambda) \text{exp}[-(w - \mu)^2/2\sigma^2_2], \;\;\; w \geq \mu
    \end{cases}
\end{equation}
For modeling in this manuscript, we use a slightly reparameterized version of 
the function in (\ref{eq:asg_function_rep}), where $\eta = \nu - \lambda > 0$. 
This constraint guarantees that the function has a peak greater than $\lambda$.

\begin{equation}
    \label{eq:asg_function_rep}
    ASG_{\theta}(w) = 
    \begin{cases}
        \lambda + \eta \text{exp}[-(w - \mu)^2/2\sigma^2_1], \;\;\; w < \mu \\
        \lambda + \eta \text{exp}[-(w - \mu)^2/2\sigma^2_2], \;\;\; w \geq \mu
    \end{cases}
\end{equation}

\subsection{Model discrepancy}

The SIR and ASG functions are useful for capturing the main trend of the ILI 
data, but as Osthsu et al. \cite[]{osthus2019dynamic} point out there may be 
systematic behavior in the ILI that these or other possible functions may not 
capture. As noted in section \ref{sec:data}, figure \ref{fig:us_ili} shows a 
regular peak at week 22 of the flu season. 
Figures \ref{fig:asg_fits} and \ref{fig:discrepancy} are used together to 
illustrate the systematic discrepancy from a fitted function. Figure 
\ref{fig:asg_fits} shows the US ILI percentage for all flu seasons from 2010 
to 2022 excluding 2020 with a best fit ASG function plotted over the ILI. The 
fits for each season were made by obtaining the maximum likelihood estimate 
(MLE) of a model given the ILI data where we assume the ASG function is the 
mean parameter of a Beta distributed random variable. Figure 
\ref{fig:discrepancy} shows the discrepancy between the model fit and the data 
for the same seasons. The grey lines show the difference between the data and 
the functions from figure \ref{fig:asg_fits} for each season, and the black 
line is the average by week over all seasons. The lines show that the ASG 
function typically underpredicts week 22 and overpredicts week 23. Perhaps 
for other weeks, like week 30 for example, there also tends to be systematic 
behavior not captured by the ASG function.

\begin{figure}[hbt!]
    \centering
    \includegraphics[scale=.5]{Images/asg_fits.png}
    \caption{Observed US national influenza-like illness (ILI) percentage for 
    seasons 2010 to 2022 excluding 2020 (grey) overlaid with MLE of an 
    asymmetric Gaussian (ASG) model for the ILI data (black)}
    \label{fig:asg_fits}
\end{figure}

\begin{figure}[hbt!]
    \centering
    \includegraphics[scale=.45]{Images/discrepancy.png}
    \caption{Difference between observed US national influenza-like illness 
    (ILI) and MLE fits for an asymmetric Gaussian (ASG) model for each season 
    2010 to 2022 excluding 2020 (grey) and the average difference of all 
    seasons (black)}
    \label{fig:discrepancy}
\end{figure}

Modeling discrepancy has been used in uncertainty analysis of 
simulators to capture systematic differences between mathematical models and 
reality \cite[]{ma2022multifidelity,brynjarsdottir2014learning,
arendt2012improving,kennedy2001bayesian}. Modeling discrepancy can lead to 
overfitting, particularly in forecasting scenarios, and may also lead to 
identifiability issues. 
Thus, care must be taken in setting parameter constraints as well as in the 
selection of prior distributions \cite[]{osthus2019dynamic,
brynjarsdottir2014learning}. 
Modeling of the discrepancy for ILI was done by Osthus et al. during the 2015 
and 2016 flu seasons where their model outperformed all others in the CDC flu 
forecasting challenge \cite[]{osthus2019dynamic}.
The terms $\gamma_w$ and $\upsilon_{s,w}$, where $s$ is the season and $w$ is the 
season week, are included in 
(\ref{eq:ili_model}) to capture the per week discrepancy between ILI and the 
function. Here $\gamma_w$ is shared across all seasons, whereas $\upsilon_{s,w}$
is specific to season $s$. In fact, to assist in identifying $\gamma_w$,
$\upsilon_{s,w}$ is only included for the current season being forecast and none of
the previous seasons.
 As in Osthus's model, 
$\gamma_w$ and $\upsilon_{s,w}$ are modeled as reverse random walks, as shown 
in (\ref{eq:rw}). 
\begin{equation}
    \begin{aligned}
    \label{eq:rw}
        \gamma_w|\gamma_{w + 1} &\overset{ind}{\sim} N(\gamma_{w+1},\sigma^2_{\gamma}), \quad
        \gamma_{W} \sim N(0,\sigma^2_{\gamma_W}) \\
        \upsilon_{s,w}|\upsilon_{s, w + 1} &\overset{ind}{\sim} N(\upsilon_{s, w+1},\sigma^2_{\upsilon})
    \end{aligned}
\end{equation}

The idea for using the reverse random walk is that there are several previous 
seasons of ILI data, and assuming the random walk captures systematic behavior, 
fitting it hierarchically over seasons can assist in predicting future behavior 
in the current season. 
Reverse random walks have also been used with success in election forecasting 
and other flu forecasting models \cite[]{osthus2021multiscale, 
osthus2019dynamic, linzer2013dynamic}. The sum to zero constraint 
$-\gamma_1 = \sum_{w=2}^W \gamma_w$ is imposed on (\ref{eq:rw}) to improve 
identifiability. An additional constraint is $\upsilon_{s, W} = 0$.

\subsection{Hospitalization model} \label{sec:hospital_model}

The second component for forecast modeling is based on the relationship between 
hospitalizations and ILI and is defined in (\ref{eq:final_hosp}).
This is an example of an autoregressive model with exogenous variables where 
the autoregressive lag is one (ARX(1)) 
\cite[]{raftery2010online,ljung1987system}.

\begin{equation}
    \begin{aligned}
    \label{eq:final_hosp}
    H_{s,w} &= \alpha_{0s} + \alpha_{1s} (ILI_{s,w} \times P) + \alpha_{2s} 
    (ILI_{s,w} \times P)^2 + \phi H_{s,w-1} + \epsilon_{s,w}\\ 
    \epsilon_{s,w} &\overset{iid}{\sim} D_s(0, \sigma_{\epsilon_s}^2 \times P, 
    \omega_s) %N(0, \sigma_{\epsilon}^2)
    \end{aligned}
\end{equation}

Figure \ref{fig:lag_scatter} helps
illustrate why this modeling decision was made.
shows scatterplots of ILI percentage and hospitalizations on the top row.
The bottom row shows scatterplots of the difference between hospitalizations and 
scaled 1 week 
lags by ILI percentage at the US national level. The plots in the right column
are from the logarithm of hospitalizations. The lines through the points is
drawn by connecting hospitalizations for consecutive weeks during the flu 
season.
These plots show that a linear or quadratic relationship between ILI percentage 
and hospitalizations does not capture much of the temporal correlation in 
hospitalizations as shown by the obvious loops drawn in the plot. 
However, when including a 
single lag of hospitalizations in the model, much of the time correlation is
accounted for as shown by the tighter and more frequent crossing of lines shown
in the bottom row. Higher than order 1 autoregressions showed little or no 
improvement over order 1, hence we consider the ARX(1) reasonably well 
suited for modeling ILI and hospitalzations.

\begin{figure}[hbt!]
    \includegraphics[scale=.2]{Images/ili_v_svat_vs_lagdiff.png}
    \caption{Scatterplots of national hospitalizations (left) or log 
    hospitalizations (right) lag differences by ILI \% where the lag is 
    scaled by $\phi$. Points are colored by season. For hospitalizations 
    $\phi = 0.6$, and for log hospitalizations $\phi = 0.77$. }
    \label{fig:lag_scatter}
\end{figure}

In model (\ref{eq:final_hosp}), $H_{s,w}$ is the number of hospitalizations for 
week $w$ in season $s$, $\epsilon_{s,w}$ is an error term distributed 
according to some distribution $D_s$ with mean parameter 0, scale parameter 
$\sigma_{\epsilon_s}$, and the additional parameter $\omega_s$ is the degrees 
of freedom parameter when $D_s$ belongs to the location-scale t (LST) family.

This model is for any location, and for fitting purposes $ILI_{s,w}$ is always 
multiplied by $P$ which is proportional to the population of the state or 
territory, in this case the total population divided by 50,000. This is done 
as a means of scaling so that the prior distribution assigned to  
$\boldsymbol{\alpha}_s = (\alpha_{0s}, \alpha_{1s}, \alpha_{2s})$, 
$\sigma_{\epsilon_s}$, and $\omega_s$ might reasonably be the same for all 
states.

Like the ILI model, the hospitalization model in (\ref{eq:final_hosp}) is also 
fit via Bayesian posterior updating. To obtain forecasts for $H_{s, w^* + i}$, 
the ILI posterior predictive distribution is used along with the posterior 
distribution for the parameters in (\ref{eq:final_hosp}). We considered three 
scenarios for model (\ref{eq:final_hosp}). A model where $D_s$ belongs to the 
normal family (NORM), $D_s$ belongs to the LST family, and one where $H_{s,w}$ 
is replaced with $\text{log}(H_{s,w} + 1)$ and $D_s$ is from a normal family, 
or $H_{s,w} + 1$ is lognormally distributed (LNORM). The population value $P$ 
is excluded from the LNORM model, and we set 
$\boldsymbol{\alpha}_{22} = \boldsymbol{\alpha}_{23}$ to help with fitting. In 
the LNORM model, if the linear parameters are not set the same for seasons 22 
and 23, the final variance was more prone to be extreme. Besides varying the 
distribution family of hospitalizations, we also considered $\alpha_{2s} = 0$ 
or there is no quadratic ILI term.

\subsection{Prior selection}

The priors selected for the ILI data model under both the SIR and ASG models 
largely follow the prior selections in Osthus et al. \cite[]{osthus2019dynamic} 
and Ulloa \cite[]{ulloa2019} with a few exceptions where changes improved 
numerical stability and/or we felt the adjusted prior made more sense for the 
problem. For model (\ref{eq:ili_model}), parameters which are common even when 
using different functions of $f_{\theta_s}(w)$ are $\kappa_s$, 
$\sigma_{\gamma}^2$, and $\sigma_{\gamma_W}^2$. For the SIR function 
$\theta_s = (S_{0s}, I_{0s}, R_{0s}, \alpha_s, \rho_s)$, and for the ASG 
function $\theta_s = (\lambda_s, \eta_s, \mu_s, \sigma_{1s}^2, \sigma_{2s}^2)$. 
For the hospitalization model in (\ref{eq:final_hosp}) the parameter to be 
estimated is $\Psi = (\alpha_{0s}, \alpha_{1s}, \alpha_{2s}, \phi, \sigma_{\epsilon_s}, \omega_s)$.

The priors assigned were mostly noninformative, though in certain cases the 
prior distributions were selected for numerical stability as was the case 
for $\sigma_{\gamma}^2$ and $\sigma_{\gamma_W}^2$. For these two scale 
parameters only, rather than assigning a half-normal prior to the standard 
deviation parameters, as recommended by Gelman \cite[]{gelman2006prior}, the 
priors were assigned to the variance parameters. Univariate parameters were 
assigned either a normal distribution prior if the support is on $\mathbb{R}$, 
a half-normal prior if the support is nonnegative, or a truncated-normal 
prior to match a more specific support. Under the ASG model, $\theta_s$ is 
modeled hierarchically over seasons so that for each season the transformed 
parameter $T(\theta_s) \sim N(\theta, \Sigma$) and priors distributions are 
assigned to $\theta$ and $\Sigma$.

Additional prior constraints were made to improve parameter identifiability. 
In Osthus et al. \cite[]{osthus2019dynamic} the initial value of the 
susceptible population compartment of the SIR model was set to $S_0 = 0.9$. 
The parameters $I_{0s}$, $\beta_s$, and $\rho_s$ were assigned informative 
priors. 
% To improve identifiability when $f_{\theta_s}(w) = ASG_{\theta_s}(w)$ 
% in (\ref{eq:ili_model}) we followed a modular Bayesian approach for fitting the 
% parameters. A modular Bayesian approach involves multiple steps of parameter 
% fitting where some parameters may be estimated without priors via maximum 
% likelihood estimation or other means. Fitting the rest of the model parameters 
% involves assigning priors and conditioning on the previously fit parameters. 
% This has been done in computer modeling to improve identifiability and other 
% issues, though Liu et al. \cite[]{liu2009modularization} warn this approach is 
% not probabilistically sound if parameter inference is a priority 
% \cite[]{jiang2015surrogate, arendt2012improving, liu2009modularization}. 
% We carried out the modular fit by first estimating the maximum likelihood 
% estimate (MLE) for the parameter $\lambda_s$ for each season in 
% (\ref{eq:asg_function_rep}) and plugging in the MLE as a fixed value.

\subsection{Hospitalization Forecasts Using the Bayesian Cut}

Rather than combining the ILI model in equation (\ref{eq:ili_model}) with the 
hospitalization model in equation (\ref{eq:final_hosp}) and computing a fully Bayesian
posterior distribution, both models were fit separately and 1 to 4 week ahead
ILI forecasts from model (\ref{eq:ili_model}) were combined with the posterior
predictive distribution of model (\ref{eq:final_hosp}) to obtain 1 to 4 week
ahead hospitalization forecasts. In this scheme, the posterior distribution of 
model (\ref{eq:ili_model}) parameters was not directly influenced by the 
hospitalization data and the posterior distribution of model 
(\ref{eq:final_hosp}) parameters was not directly influenced by model 
(\ref{eq:ili_model}) parameters. Hence there was limited feedback between the
two components, but propagation of uncertainty in ILI forecasts was still 
permitted in the estimating of hospitalization forecasts.

This two-component modeling is remeniscent of the "Bayesian cut" as 
described in \cite{plummer2015cuts} and \cite{nott2023bayesian}. Cutting a model
into two component models or "modules" may be justfied for several reasons
including to reduce time to fit the model, avoiding mixing issues of posterior
sampling distributions, improving predictive performance or to prevent
model misspecification in one model to influence the other 
\cite[]{nott2023bayesian, jacob2020unbiased, jacob2017better, plummer2015cuts}.

The decision to cut the model into two components was made for several reasons.
When fitting the fully Bayesian joint model, the time to fit a single model using
Markov chain Monte Carlo (MCMC) sampling and
obtain forecasts was sometimes upwards of 4 hours whereas fitting the two models
separately usually took less than 1 hour. When assessing convergence of model
parameters, it was rare that posterior sampling chains for the joint model
showed convergence even after 55,000 iterations, whereas the component model 
fits gave no signs to be concerened with convergence. And most importantly,
the forecasts from the component models generally outperformed those of the 
joing model in terms of minimizing proper scoring rules. 
The process for obtaining forecasts is described in the following paragraphs 
and computation is described in section \ref{sec:implementation_posterior}.

Model (\ref{eq:ili_model}) is fit via Bayesian posterior updating. Future ILI 
forecasts are obtained via the posterior predictive distribution where for
week $w$, the predictive distribution is obtained by integrating over the 
parameters $\pi_s$, $\kappa_s$, $\sigma^2_{\gamma}$, and $\sigma^2_{\gamma_W}$ 
as in (\ref{eq:ili_post}) where 
$p(\boldsymbol{\pi}_s, \kappa_s, \sigma^2_{\gamma}, \sigma^2_{\gamma_W} | \textbf{ILI})$ 
is the density function of the posterior distribution for the model parameters. 
If the current week is $w$ then the desired forecasts are for weeks $w + i$ 
where $i$ is a positive integer. Here $\widetilde{ILI}_{s,w + i}$ represents the 
predicted or forecast ILI at $i$ weeks in the future.

\begin{equation}
    \label{eq:ili_post}
    p(\widetilde{ILI}_{s,w +i} | \textbf{ILI}) = \int \int \int \int 
    p(\widetilde{ILI}_{s,w + i} | \boldsymbol{\pi}_s, \kappa_s, 
    \sigma^2_{\gamma}, \sigma^2_{\gamma_W}) p(\boldsymbol{\pi}_s, \kappa_s, 
    \sigma^2_{\gamma}, \sigma^2_{\gamma_W} | \textbf{ILI}) 
    d\boldsymbol{\pi}_s d \kappa_s d \sigma^2_{\gamma} d \sigma^2_{\gamma_W}
\end{equation}
% \jarad{It doesn't make sense to have 4 $\int$ here since $\pi_s$ is a vector.}
% \spencer{Isn't that captured in $d \pi_s$?}
Hospitalization forecasts $\widetilde{H}_{s,w + i}$ at week $w + i$ are then
obtained by combining the posterior predictive distribution of model 
(\ref{eq:final_hosp}) with (\ref{eq:ili_post}) and integrating over 
$\widetilde{ILI}_{s,w + i}$ as in (\ref{eq:hosp_post_forc}). Here 
$\boldsymbol{\alpha}$ represents all regression and variance parameters in
model (\ref{eq:final_hosp}).

\begin{equation}
\begin{aligned}
\label{eq:hosp_post_forc}
    p(\tilde{H}_{s,w + i} | \textbf{ILI}, \textbf{H}, \boldsymbol{\alpha}) &= 
       \int p(\tilde{H}_{s,w + i} | \textbf{H}, \textbf{ILI}, 
       \widetilde{ILI}_{s,w + i}, \boldsymbol{\alpha}) 
       p(\widetilde{ILI}_{s,w + i} | \textbf{ILI}) d\widetilde{ILI}_{s,w + i}
\end{aligned}
\end{equation}

\subsection{Parameter estimation and posterior predictive sampling}
\label{sec:implementation_posterior}

The models were fit via MCMC sampling using the 
\texttt{cmdstanr} package which was developed and is maintained by the Stan 
Development Team \cite[]{stan2024manual} \cite[]{gabry2022stan}. Stan 
implements Hamiltonian Monte Carlo (HMC) sampling with the No-U-turn 
sampler \cite[]{hoffman2014no}. The \texttt{cmdstanr} package provides 
several diagnostic statistics for assessing the sampler.
As mentioned, most model parameter prior distributions were intended to be 
uninformative.
Plots of posterior distributions for select parameters from ILI and 
hospitalization models are shown in the supplementary material 
\cite[]{wadsworth2024bas}.

We assessed the model fit for four ILI models. These included the SIR and 
ASG models and models with and without discrepancy modeling. When discrepancy 
is included, the models are denoted as SIRD and ASGD. These models were fit 
using US national data from 2010 to 2023 flu seasons, where data from the 
2020 season was excluded because of the unique behavior during that season. 
Assessment for hospitalization modeling was done for six different models. 
These include NORM, LNORM, and LST models, and models where a quadratic ILI 
term is included or excluded. To assess posterior sampling convergence, models 
were fit to data where ILI and hospitalization data up to week 14 of the 2023 
season was included. Sampling was done with four chains where from each chain 
60,000 posterior draws were sampled, and the first 10,000 draws were discarded 
as a burn-in. The $\hat{R}$ statistic \cite[]{vehtari2021rank} and the 
effective sample size ($ESS$) \cite[]{gelman2013bayesian} were calculated for 
each parameter. Tables \ref{tab:ili_diagnostics} and 
\ref{tab:hsop_diagnostics} summarize the maximum $\hat{R}$ and the minimum 
$ESS$ over all parameters for ILI and hospitalization models respectively. 
Forecast models for all other weeks of the season were fit using one chain of 
60,000 draws where the first 10,000 draws were discarded as a burn-in. 
% \jarad{How long does this take to run? 8,000 inferential iterations seems very small.}
For parameters of the ASG models that were prone to cause trouble in posterior 
sampling, the starting values were set to be the MLEs.

\begin{table}
\caption{Maximum $\hat{R}$ and minimum $ESS$ over all parameters for four ILI 
models fitted on US data for week 14 of the 2023 season}
\label{tab:ili_diagnostics}
\begin{center}
    \begin{tabular}{c|cccc}
     & ASG & ASGD & SIR & SIRD \\
     \hline
     $\hat{R}$ & $< 1.001$ & $< 1.001$ & $< 1.001$ & $1.001$ \\
     $ESS$ & 72,057 & 9,980 & 17,745 & 6,259 \\
    \end{tabular}
    \end{center}
\end{table}

\begin{table}
\caption{Maximum $\hat{R}$ and minimum $ESS$ over all parameters for six 
hospitalization models fitted on US data for week 14 of the 2023 season}
\label{tab:hsop_diagnostics}
\begin{center}
    \begin{tabular}{c|cccccc}
     & NORM & NORM$^2$ & LNORM & LNORM$^2$ &  LST &  LST$^2$ \\
     \hline
     $\hat{R}$ & $< 1.001$ & $< 1.001$ & $< 1.001$ & $< 1.001$ & $< 1.001$ & 
     $< 1.001$\\
     $ESS$  & 71,266 & 73,554 & 46,747 & 54,975 & 7,660 & 62,663 \\
    \end{tabular}
    \end{center}
\end{table}

To obtain forecast distributions of hospitalizations, draws from the posterior 
predictive distribution from the ILI model were used in conjunction with the 
posterior distribution of the hospitalizations model. When fitting model 
(\ref{eq:ili_post}), MCMC samples of $\widetilde{ILI}_{s,w:(w + 4)}$ were 
saved. Model (\ref{eq:final_hosp}) was fit and MCMC samples for the marginal 
distributions for the model parameters were saved. To obtain forecast 
distributions for $H_{s,w + i}$ where $i \in \{1,2,3,4\}$, the following 
steps are repeated $K$ times where $K$ is an integer for the number of desired 
samples. We set $K = 50,000$.

\begin{quote}
\begin{enumerate}[Step 1:]
  \item Sample $\widetilde{ILI}^*_{s,w:(w + 4)}$
  \item Sample $\alpha_{0s}^*$, $\alpha_{1s}^*$, $\alpha_{2s}^*$, $\phi^*$, 
  $\sigma^*_{\epsilon_s}$, $\omega^*_s$ from respective marginal posterior 
  distributions
  \item Sample $H^*_{s,w + i}$ from $D(\omega_s^*, \mu_{s, w + i}^*,\sigma^2_{\epsilon_s}$), where
  \begin{itemize}
    \item[] $\mu_{s,w + i}^* = \alpha_{0s}^* + \alpha_{1s}^* (ILI_{s,w + i}^* \times P) + \alpha_{2s}^* (ILI_{s,w + i}^* \times P)^2 + \phi^* H^*_{s,w + i - 1}$
  \end{itemize}
   \item Repeat step 3 for $i \in \{1,2,3,4\}$ to obtain $H^*_{s,(w + 1):(w + 4)}$
  \item Repeat steps 1-4 $K$ times
\end{enumerate}
\end{quote}
The sample $\{H^*_{s,w + i}\}^K$ was then used as the probabilistic forecast 
for hospitalizations at week $w + i$. For the forecast competition analysis in 
section \ref{sec:analysis}, all negative values of $\{H^*_{s,w + i}\}^K$ were 
set to 0 to reflect realistic values of hospitalizations and comply with the 
FluSight forecasting rules.

\section{Simulation Study} \label{sec:simulation2}
In this section, we present a simulation study conducted for comparing ILI 
models and further assessing the hospitalization forecast model. US ILI data 
is used, and hospitalization data is simulated. 
A leave-one-season-out (LOSO) approach was combined with a Monte Carlo 
simulation approach. For each replication, we simulated log-hospitalizations 
for all weeks during seasons 2010, ..., 2022, excluding 2020, using the 
existing ILI data as a predictive covariate. Each season was in turn 
``left-out" and treated as if it was the most recent season which we desired 
to forecast. Fitting and forecasting was then done for weeks 14, 20, 26, 32, 
and 38 of the left out season, giving two weeks that tend to occur as flu 
cases increase, two as cases decrease, and one that occurs when the cases may 
be increasing or decreasing. Week 20 is a week leading up to the holiday week 
22 where ILI typically has a local peak. We were particularly interested in 
how important modeling discrepancy is for forecasting at week 20.

Besides the 4 ILI models, we also included forecasts for a naive baseline
(BASE) model and an autoregressive integrated moving average (ARIMA) model. 
The BASE model is a simple random walk with drift model, and the ARIMA model
is fit by selecting the best possible model give the data according to a 
fitting criterion. Both the BASE and ARIMA ILI models were fit using the 
\texttt{forecast} package in \texttt{R} \cite{hyndman2008forecast}. Samples
were then drawn from the forecast distributions, and these were then used
along with the model in (\ref{eq:final_hosp}) to make hospitalization 
forecast distributions.

For the simulation of hospitalizations, the parameters 
$\boldsymbol{\alpha}_s = (\alpha_{0s}, \alpha_{1s}, \alpha_{2s})$ and 
$\sigma^2_{\epsilon_s}$ from the hospitalization model in 
(\ref{eq:final_hosp}) were considered the same across all seasons so that 
all $\boldsymbol{\alpha}_s = \boldsymbol{\alpha}$. 
The values for $\boldsymbol{\alpha}$, $\sigma^2_{\epsilon}$, and $\phi$ were 
estimated by fitting model (\ref{eq:final_hosp}) using ILI and hospitalization 
data from the 2022 season. For fitting, the hospitalization data was first 
log-transformed. We took $\boldsymbol{\alpha}_{\phi} = (\boldsymbol{\alpha}, \phi)$
and assigned the noninformative prior 
$p(\boldsymbol{\alpha}_{\phi}, \sigma^2_{\epsilon}) \propto 1/\sigma^2_{\epsilon}$. 
The marginal posterior distribution 
$\boldsymbol{\alpha}_{\phi} | \sigma^2_{\epsilon}, \boldsymbol{H_{22}}$ 
was then the established posterior multivariate normal distribution and 
$\sigma^2_{\epsilon} | \boldsymbol{H_{22}}$ the inverse-$\chi^2$ posterior 
distribution \cite[]{gelman2013bayesian}. The posterior means of those 
parameters were used as the values from which log-hospitalizations were 
simulated. The number of replicates in the simulation was 500.

Model comparison was done by calculating a proper scoring rule for each 
forecast.
Proper scoring rules are the 
current standard for comparing performance between probabilistic forecasts 
and selecting the best forecasts according to the notion of maximizing 
sharpness subject to (auto-)calibration 
\cite[]{gneiting2007probabilistic, tsyplakov2013evaluation}. Proper scoring 
rules are commonly used in forecast comparison and have the property that a 
forecaster is incentivized to be honest in the reporting of their forecasts 
\cite[]{gneiting2007strictly, gneiting2014probabilistic}.
The proper scoring rules primarily used in the FluSight competition and the 
COVID-19 Forecast Hub and which we will use herein are the 
weighted interval score (WIS) and the log-weighted interval score (LWIS).
\cite[]{mathis2024evaluation, bracher2021evaluating}.
The WIS is used for scoring quantile or interval 
forecasts 
\cite[]{gneiting2007strictly, gneiting2014probabilistic, bracher2021evaluating} 
and is defined in (\ref{eq:wis2}) where $Q$ is a forecast represented by all 
included quantiles, $B$ is the number of intervals,  $y^*$ is the observed 
value targeted by the forecast, $w_0 = 1/2$ and $w_b = \alpha_b / 2$ are 
weights for each interval, and $\alpha_b$ is the nominal level of the $b^{th}$ 
interval. $IS_{\alpha}$ is the interval score (IS), a proper scoring rule for a 
single interval as defined in (\ref{eq:is}). Here 
$\vmathbb{1}\{\cdot\}$ is the indicator function.
The LWIS is the same as the WIS except that it is evaluated 
over the log of quantiles and the log of the observed value.

\begin{equation}
\label{eq:wis2}
        WIS_{0,B}(Q, y^*) = \frac{1}{B + 1/2} \times (w_0\times |y^* - median| + \sum_{b=1}^B \{w_k \times IS_{\alpha_b}(Q, y^*) \} )
\end{equation}

\begin{equation}
\label{eq:is}
        IS_{\alpha}(l,r;y^*) = (r-l) + \frac{2}{\alpha}(l - y^*)\vmathbb{1}\{y^* < l\} + \frac{2}{\alpha}(y^* - r) \vmathbb{1}\{y^* > r\}
\end{equation}

%  The CRPS is defined 
% in (\ref{eq:crps2}). Here $F(\cdot)$ is the forecast distribution function, 
% and $y^*$ is the observed targeted value of the forecast. The orientation of 
% the CRPS is negative, meaning the smaller the score the better. Here 
% $\vmathbb{1}\{\cdot\}$ is the indicator function.
% The CRPS is calculated using a quantile decomposition from 
% \cite[]{laio2007verification}. 
% 
% \begin{equation}
%     \text{CRPS}(F, y^*) = \int_{-\infty}^{\infty} (F(x) - \vmathbb{1}\{y^* \leq x\})^2 dx
%     \label{eq:crps2}
% \end{equation}

% \begin{equation}
%     \text{LogS}(f, y^*) = -\text{log}(f(y^*))
%     \label{eq:logs2}
% \end{equation}
% Both the CRPS and LogS are calculated using the \texttt{scoringRules} package in \texttt{R} \cite[]{jordan2019scoring}.
% To calculate the LogS when given MCMC samples from a posterior predictive distribution, a continuous density function was first estimated via kernel density estimation.

The forecasts used in this section and in the subsequent section follow the 
FluSight convention of being quantiles corresponding to the 23 given 
probabilities $\\(0.010, 0.025, 0.050, 0.100, 0.150, …, 0.950, 0.975, 0.990)$
making each forecast a set of 11 predictive intervals and a predictive median.
Figure \ref{fig:wis_by_week_horizon} shows 
boxplots of the WIS for the 500 simulation replicates faceted by week 
on the y-axis and horizon on the x-axis. 
% Note that in figure \ref{fig:logs_by_week_horizon}, only LogS values of less than 17 are included to improve visualization.
In most plots the ASGD model shows the best performance, 
especially during week 20 where
forecasts are near weeks 22 and 23 which show a seasonal peak and trough in 
figure \ref{fig:discrepancy}. The SIRD also shows improved performance over the
SIR model, though its performance does not match that of the ASG or ASGD.
Here the ASGD model regularly outperforms
the BASE and ARIMA models. The three other nonlinear models do not perform as
well.
The ASG model
tends to outperform the SIR and SIRD models suggesting that the additional 
flexibility of the five parameters ASG function 
is often better for forecasting than that of the three parameter SIR function.
The model with the best performance is the ASGD.The ASGD model shows 
especially good forecast skill at week 20 around where we would expect the 
discrepancy to be most useful.

% Notable plots here are for week 20 where the two models accounting for the 
% ILI discrepancy, ASGD and SIRD, are the best performing forecast models, with 
% lower medians and lower variation than the two models not accounting for ILI 
% discrepancy. This makes sense given the forecasts at week 20 are forecasting 
% weeks, 22 and 23 where figure \ref{fig:discrepancy} shows a seasonal peak and 
% trough. This confirms that there is value in modeling discrepancy at least 
% during around the holiday weeks near week 22. Among the four models, ASG 
% tends to have the lowest CRPS values though this is not always the case. 
% This suggests that the additional flexibility of the five parameter ASG 
% function is often better for forecasting than that of the three parameter SIR 
% function. 
% Additional
% plots in the supplemtary material \cite[]{wadsworth2024bas} further show that 
% the ASG models often but not always outperform the SIR models according to the 
% CRPS and another common proper scoring rule known as the logarithmic score.

\begin{figure}[hbt!]
    \centering
    \includegraphics[scale=.57]{Images/wis_by_week_horizon.png}
    \caption{Boxplots of weighted interval score (WIS) for the 
    four Bayesian nonlinear ILI models, a baseline model (BASE) and
    an autoregressive integrated moving average model (ARIMA) 
    over all seasons in the simulation study faceted by horizon 
    (x-axis) and season week (y-axis). Horizons include 1-4 week ahead forecasts 
    and weeks include weeks 14, 20, 26, 32, and 38 of the flu season}
    \label{fig:wis_by_week_horizon}
\end{figure}

\begin{table}[hbt!]
\centering
\begin{tabular}{lrrrr}
Model & WIS & MAE & 50\%Coverage  & 95\% Coverage \\ 
  \hline
ASGD & 0.09 & 0.15 & 42 & 91 \\ 
  ASG & 0.13 & 0.18 & 33 & 71 \\ 
  ARIMA & 0.13 & 0.18 & 35 & 68 \\ 
  BASE & 0.15 & 0.20 & 33 & 65 \\ 
  SIRD & 0.15 & 0.23 & 33 & 78 \\ 
  SIR & 0.21 & 0.29 & 20 & 51 \\ 
\end{tabular}
\end{table}

\section{Analysis of forecasts for 2023 flu season}
\label{sec:analysis}

In this section we apply the forecast models to make forecasts for the 2023 
flu season weekly hospitalizations and compare the results with those of 
forecasts submitted to FluSight. The scoring of the forecasts is in the 
context of the FluSight competition where each competing forecast was 
submitted as a set of 23 quantiles corresponding to give probabilities.
Forecasts of 1, 2, 3, and 4-week ahead hospitalization counts were requested, 
and forecasts were made at the state and national level. The first week of 
forecasting took place during the week of October 7, 2023, and the final week 
was the week of May 4, 2024 making 30 total weeks of forecasts.
 The same format was used during the 2021 and 2022 seasons and for the 
 COVID-19 Forecast Hub \cite[]{mathis2024evaluation, bracher2021evaluating}. 
% Nearly teams participated and forecasts from over 40 models were submitted throughout the duration of the season. 
 Primary scores for evaluating each forecast were the WIS and LWIS

The prior distributions under the SIR model are in (\ref{eq:sir_prior}). 
Because we set $S_{0s} = 0.9$, prior distributions were assigned only to 
$I_{0s}$, $\beta_s$ and $\rho_s$, recalling the parameter for the recovery 
rate $\delta_s = \rho_s \beta_s$. Here $\vmathbb{1}_{A}$ represents the 
indicator function for values within the set $A$.

\begin{equation}
\begin{aligned}
    \label{eq:sir_prior}
        I_{0s} &\sim N(0.005, 0.03) \vmathbb{1}_{(0, .1)} \\
        \beta_s &\sim N^+(0.8, 0.3) \\
        \rho_s &\sim N^+(0.68, 0.08)
\end{aligned}
\end{equation}
For the ASG model, the MLE 
$\hat{\theta}_s = (\hat{\lambda}_s, \hat{\eta}_s, \hat{\mu}_s, \hat{\sigma}_{1s}^2, \hat{\sigma}_{2s}^2)$ 
was calculated. % and $\hat{\lambda}_s$ was accepted as a fixed value. 
These estimates were used as starting values for posterior sampling. 
The transformation 
$T(\theta_s) = (\lambda_s, \text{log}(\eta_s), \mu_s, \text{log}(\sigma_{1s}^2), \text{log}(\sigma_{2s}^2))$ 
was made, and a prior distribution was assigned to $T(\theta_s)$.
The prior distributions for the ILI model under the ASG function are shown 
in (\ref{eq:ili_prior}). These are slightly informative priors because for 
most parameters we have an idea what reasonable values may be. Here 
$m = (-4.5, 0.3, 23, 3.69, 4.7)$ and $C = \text{diag}(0.3, 0.2, 5, 2, 2)$
where $\text{diag}(\cdot)$ 
is the diagonal matrix for the given entries.

\begin{equation}
\begin{aligned}
\label{eq:ili_prior}
                T(\theta_s) &\overset{ind}{\sim} MVN(\theta, \Sigma) \\
                T(\theta) &\overset{ind}{\sim} MVN(m, C)\\
                \Sigma &= \text{diag}(\zeta^2_1,...,\zeta^2_4) \\
                \zeta_i &\overset{ind}{\sim} N^+(0,4^2)
\end{aligned}
\end{equation}
Parameters shared by both the SIR and ASG models are the scale parameter 
$\kappa_s$ and the discrepancy parameters $\sigma_{\gamma}^2$ and 
$\sigma_{\gamma_W}^2$. These priors are in (\ref{eq:shared_ili_priors}).

\begin{equation}
\begin{aligned}
    \label{eq:shared_ili_priors}
        \kappa_s &\overset{ind}{\sim} N^+(0, 10,000^2) \\
        \sigma_{\gamma}^2 &\sim N^+(0, .02^2) \\
        \sigma_{\gamma_W}^2 &\sim N^+(\hat{\sigma}_W^2, 1^2) \\
        \sigma_{\upsilon_{s,w}} &\sim N^+(0, .1^2)
\end{aligned}
\end{equation}
Because of the limited information for estimating $\sigma_{\gamma_W}^2$, 
we selected an informative prior distribution by first estimating  
$\hat{\sigma}_{\gamma_W}^2$. This was estimated for a given state by first 
calculating the MLE for $\theta_s$. 
% We set the likelihood function as $\mathcal{L}(\theta_s) = \prod_{w = 1}^W f_{\theta_s}(w)$ and  calculated $\hat{\theta}_s$ where $\label{eq:theta_hat}
%     \hat{\theta}_s = \underset{\theta_s}{\mathrm{argmin}} \, \mathcal{L}(\theta_s)$.
$\widehat{ILI}_{s,w}$ was then predicted such that 
$\text{logit}(\widehat{ILI}_{s,W}) = f_{\hat{\theta}_s}(W)$ for each season. 
Then $\hat{\sigma}_{\gamma_W}^2$ was calculated as the estimated variance over 
seasons of the differences 
$\text{logit}(\widehat{ILI}_{s,w}) - \text{logit}(ILI_{s,w})$. 
When $f_{\theta_s}(w)$ was the SIR function, $\widehat{\theta}_s$ was 
calculated using the \texttt{mle2} function in the \texttt{bblme} package 
\cite[]{bolker2023bblme} and the \texttt{ode} function in the \texttt{deSolve} 
package \cite[]{soetaert2010desolve} function in \texttt{R}. Where the ASG 
function was used,  $\hat{\theta}_s$ was calculated using the \texttt{optim} 
function.
The prior for the hospitalization models are in (\ref{eq:shared_hosp_priors}).

\begin{equation}
\begin{aligned}
\label{eq:shared_hosp_priors}
        \alpha_{0s} &\overset{ind}{\sim} N(0, 5^2)\\
        \alpha_{1s} &\overset{ind}{\sim} N(0, 5^2)\\
        \alpha_{2s} &\overset{ind}{\sim} N(0, 5^2) \\
        \phi &\sim N(0, .4)\vmathbb{1}_{(-1,1)} \\
        \sigma_{\epsilon_s} &\overset{ind}{\sim} N^+(0, 4^2) \\
        \omega_s &\overset{ind}{\sim} N^+(0, 15^2)
\end{aligned}
\end{equation}

We fit 24 forecast models for each location for all 30 weeks, and for each 
week forecast 1-4 week ahead hospitalizations. The 24 models included all 
combinations of ASG, ASGD, SIR, and SIRD ILI models, the NORM, LNORM, and LST 
hospitalization models, and both quadratic and linear hospitalization models.
Figure \ref{fig:normal_flu_forecasts} shows 1-4-week ahead forecasts for US 
flu hospitalizations during the 2023 season under the NORM hospitalization 
model. The forecasts shown are for weeks 14, 20, 26, and 32. The predictive 
bands are the 50\% and 95\% predictive intervals.
These plots show a tendency to often underpredict hospitalizations. The 
quadratic models appear to do a better job predicting  hospitalizations at 
the season peak but a poorer job predicting after the peak, whereas the linear 
models seem to predict well or slightly overpredict after the peak.

Of the 24 models, the one which produced the best forecasts was that using
the ASGD ILI model and the NORM quadratic hospitalization model. The results of
that analysis may be found in the supplementary materials ??. For the remainder 
of this section, forecast performance from the best model, which we denote
as ASG\_NORM2, is compared to that of forecasts from 20 models from which
submissions were made to FluSight in 2023. 
The models selected are those which were not ensemble models. Among the models
was the FluSight baseline. The baseline forecast had as a median the most 
recently observed 
hospitalization count, and the forecast uncertainty was based on 
differences between 
previous hospitalizations. The baseline forecasts was
similar to the baseline forecast used in
previous flu forecasting seasons and in the COVID-19 hub 
\cite[]{mathis2024evaluation, Cramer2022-hub-dataset}.

Figure \ref{fig:lwis_cover_sum} shows forecast results for 2023 flu 
hospitalization forecasts for all 21 models being compared but with results
for the ASG\_NORM2 and FluSight baseline models highlighted. 
For each week of the 2023 season, the  
LWIS averaged over location and horizon 
for each model is shown in a). Boxplots of LWIS for 1, 2, 3, and 4-week ahead
horizons for the ASG\_NORM2 and baseline models are shown in b). And 
overall empirical coverage of the 11 predictive intervals for all models is 
shown in c) with ASG\_NORM2 and the baseline model again highlighted. 
These plots show that over the season the ASG\_NORM2 model tended to 
outperform most others in terms of LWIS, and in general is superior to the 
baseline model. The coverage plot shows that the ASG\_NORM2 model is well 
calibrated at most levels, with slight overcoverage at the lower levels
up to about 80\%, and has better coverage than all other models. Similar results
for the WIS are shown in the supplementary materials ??.

\begin{figure}[hbt!]

    % \centering
    \includegraphics[scale = .6]{Images/lwis_cover_sum.png}
    \caption{a) Averaged over location and horizon log-weighted interval score
    (LWIS) for each week during the season for the 21 models being compared. 
    Each grey line represents one model, but the ASG\_NORM2 (blue) and 
    FluSight-baseline (red) are colored to stand out.
    b) Boxplots of all LWIS scores for 1, 2, 3, and 4-week ahead horizons 
    for the ASG\_NORM2 (blue) and FluSight-baseline (red) models.
    c) Plot of empirical coverage against theoretical coverage for 11 
    predictive intervals from the 10\% nominal level to the 98\%. The black
    line $x = y$ shows where optimal coverage would fall.
    Each grey line represents one model, but the ASG\_NORM2 (blue) and 
    FluSight-baseline (red) are colored to stand out.}
    \label{fig:lwis_cover_sum}
\end{figure}

Figure \ref{fig:state_and_date_lwis} shows the average LWIS for each model
relative to the ASG\_NORM2 model by location in the top tile plot and by 
season week in the bottom tile plot. That is, the value for each tile is the 
ratio of the average ASG\_NORM2 LWIS to the average for the labeled model at the
labeled location or week. A value less than 1 shows superior performance by
ASG\_NORM2 with increasing superiority as the value goes to 0. A value larger
than 1 shows inferior performance, and a value of 1 shows equal performance.
Tiles are grey where the labeled forecast model made no forecast submissions.
Viewed this way, only UGA\_flucast-INFLAenza appears to potentially outperform
ASG\_NORM2.

\begin{figure}[hbt!]
\begin{adjustwidth}{-3cm}{-1.5cm}
    % \begin{minipage}{\linewidth}
    % \centering
    \makebox[\linewidth][l]{
    \includegraphics[scale = .45]{Images/state_and_date_lwis.png}
    }
    \end{adjustwidth}
    \caption{}
    \label{fig:state_and_date_lwis}
    % \end{minipage}

\end{figure}

\begin{table}[ht]
\centering
\begin{adjustwidth}{-3cm}{-1.5cm}
\begin{tabular}{lrrrrrrrr}
  % \hline
 & MWIS & MLWIS & MedWIS & MedLWIS & MSEC & \% Forcs & RWIS & RLWIS\\ 
  \hline
ASG & 62 & 0.34 & 14 & 0.23 & 0.001 & 100 & 1 & 1 \\ 
  PSI-PROF & 63 & 0.35 & 15 & 0.23 & 0.003 & 100 & 0.90 & 0.97 \\ 
  CEPH-Rtrend\_fluH & 69 & 0.39 & 13 & 0.28 & 0.004 & 100 & 0.81 & 0.86 \\ 
  FluSight-baseline & 85 & 0.41 & 16 & 0.32 & 0.023 & 100 & 0.63 & 0.82 \\ 
  UM-DeepOutbreak & 83 & 0.46 & 21 & 0.39 & 0.010 & 99 & 0.64 & 0.72 \\ 
  cfarenewal-cfaepimlight & 76 & 0.37 & 15 & 0.26 & 0.001 & 99 & 0.72 & 0.92 \\ 
  MOBS-GLEAM\_FLUH & 66 & 0.35 & 15 & 0.22 & 0.021 & 98 & 0.87 & 0.95 \\ 
  LosAlamos\_NAU-CModel\_Flu & 242 & 0.64 & 24 & 0.42 & 0.153 & 93 & 0.21 & 0.53 \\ 
  UGuelph-CompositeCurve & 132 & 0.52 & 25 & 0.41 & 0.018 & 93 & 0.43 & 0.62 \\ 
  LUcompUncertLab-chimera & 84 & 0.43 & 17 & 0.33 & 0.004 & 92 & 0.67 & 0.71 \\ 
  UGA\_flucast-INFLAenza & 74 & 0.32 & 15 & 0.20 & 0.003 & 90 & 0.77 & 1.07 \\ 
  cfa-flumech & 101 & 0.41 & 17 & 0.27 & 0.019 & 89 & 0.58 & 0.85 \\ 
  UGA\_flucast-Copycat & 76 & 0.35 & 16 & 0.21 & 0.012 & 87 & 0.81 & 0.98 \\ 
  GT-FluFNP & 105 & 0.48 & 18 & 0.38 & 0.046 & 86 & 0.45 & 0.73 \\ 
  UNC\_IDD-InfluPaint & 107 & 0.35 & 16 & 0.24 & 0.043 & 78 & 0.58 & 0.95 \\ 
  NU\_UCSD-GLEAM\_AI\_FLUH & 98 & 0.39 & 22 & 0.24 & 0.014 & 74 & 0.68 & 0.83 \\ 
  SigSci-CREG & 59 & 0.49 & 12 & 0.31 & 0.018 & 71 & 0.62 & 0.74 \\ 
  PSI-PROF\_beta & 77 & 0.32 & 18 & 0.20 & 0.004 & 67 & 0.87 & 0.93 \\ 
  NIH-Flu\_ARIMA & 230 & 0.34 & 28 & 0.23 & 0.001 & 20 & 0.57 & 0.88 \\ 
  GH-model & 161 & 1.52 & 28 & 1.47 & 0.373 & 17 & 0.18 & 0.27 \\ 
  UGA\_flucast-OKeeffe & 27 & 0.59 & 8 & 0.43 & 0.003 & 10 & 0.60 & 0.58 \\  
\end{tabular}
\end{adjustwidth}
\end{table}

It should not be assumed that these two takeaways will apply for future flu 
seasons. As suggested by the simulation study in the previous section, it may 
often be the case that the ASG ILI model is better for forecasting. Indeed, a 
close examination of figures \ref{fig:us_lwis} and \ref{fig:lwis_by_dist_loc} 
shows that it is common for the ASG models to show better forecasting skill 
than the SIR models.

\section{Conclusion}
\label{sec:conclusion}

 In this manuscript we introduce a statistical modeling framework which allows 
 for the incorporation of several ILI forecast modeling methods. Specifically, 
 we built upon Osthus et al. \cite[]{osthus2019dynamic} and introduced a 
 framework for modeling ILI which includes the use of an arbitrary function 
 for modeling the main trajectory of ILI along with modeling the discrepancy. 
 We model flu hospitalizations by incorporating the ILI forecast model into a 
 model forecasting hospitalizations where hospitalization predictions are a 
 linear or quadratic function of ILI.

The simulation study in section \ref{sec:simulation2} suggests the ASG 
function in ILI modeling may slightly outperform the SIR model according to 
the LogS and CRPS scoring rules, but in the analysis of the 2023 flu season 
forecasts, the SIR model was overall superior. The results from both the 
simulation study and the real data analysis suggest that the addition of a 
discrepancy component in ILI modeling may improve forecasts especially near 
the holiday week between Christmas and New Years day. It should not be assumed 
that these conclusions may be generalized for all locations in the US, weeks 
of a season, or for future flu seasons. 
% Rather than deciding to use or not to use certain of the modeling techniques here, a more promising option may be to combine forecasts from various models into an ensemble forecast.

Forecasting the seasonal influenza outbreak remains a challenging task for 
forecasters. The general modeling framework in this manuscript is successful 
under diverse modeling conditions for all locations in the US and may 
contribute to future forecasting efforts. All forecast models were presented 
as separate from one another, but in the various locations and times of the 
flu season, different models perform better than others. To build on the work 
done here, a natural step forward would be to combine all or a few selected 
forecasts into an ensemble forecast. Such an ensemble may work to cancel out 
certain model biases or highlight model strengths, leading to more robust 
forecasts.