### 2025-11-28 response_round2.tex
Start responses

\documentclass[ba]{imsart}

\RequirePackage{amsthm,amsmath,amsfonts,amssymb}
\RequirePackage[numbers]{natbib}
%\RequirePackage[authoryear]{natbib}%% uncomment this for author-year citations
\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue,backref=page,backref=page]{hyperref}
\RequirePackage{graphicx}

%%%%%%%%%%%packages added by Spencer%%%%%%%%%%%%%
\usepackage{booktabs} %for the table in the analysis section
\usepackage{enumerate} %for the short algorithm
\usepackage{subfigure} %for making plots with multiple images
\usepackage{tabularx} %for the spacing in tabular
\usepackage{newtxmath} %for making indicator function
\usepackage{changepage} %for adjusting the margins
\usepackage{dsfont} %for getting the indicator function
\usepackage{mathtools}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\1}[1]{\mathds{1}\left[#1\right]}
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}

\usepackage{enumerate}
\usepackage{hyperref}
\usepackage{amsmath}

\usepackage{xcolor}
\newcommand{\jarad}[1]{{\color{red} Jarad: #1}}
\newcommand{\spencer}[1]{{\color{red} Response: #1}}

\begin{document}

\Huge
\noindent AE1
\normalsize
\\~\\

I think the authors did a great job in this newly submitted version and 
addressed most of the earlier raised comments in a nice and convincing way. 
The paper is well written and interesting to read.

I only have two remaining comments.

\begin{enumerate}[-]

  \item Section 3.7. End of of first paragraph: ``..., but propagation of 
  uncertainty   in ILI forecasts was stil permitted in the estimating of 
  hospitalisation forecasts.``

  Please clarify more explicitly how uncertainty was propagated between the two 
  model components. My understanding is that this was achieved by incorporating 
  not only point estimates of the parameters but also samples from the predictive 
  distribution of the ILI model. If this is correct, please state it clearly 
  already at this stage and consider adding a brief discussion of the advantages 
  and disadvantages of this approach.
  \\~\\
  \item I recommend including the statement raised by the reviewer in the first 
  submission—that WIS approximates CRPS (ideally supported with a reference)—and 
  therefore leads to essentially the same conclusions. I think the motivation for 
  using WIS instead of CRPS is well explained, especially given that WIS is the 
  primary evaluation measure used in the FluSight competition 
  \\~\\

  \emph{The following statements were added to page 17.}
  \begin{quote}
  \emph{
      Commonly used proper scoring rules for continuous distribution forecasts
      include the log score and the continuously
      ranked probability score (CRPS).}
      \\~\\  
  \emph{
      We note that as the number of forecast intervals $B$ increases, the WIS becomes 
      arbitrarily close to
      the CRPS. Thus, forecast scores under the CRPS and the WIS should be similar
      \cite[]{bracher2021evaluating}.}
      \\~\\
\end{quote}

\end{enumerate}

\Huge
\noindent R1
\normalsize
\\~\\
\noindent The authors have addressed all my comments from the first round. I only have two remaining points
on the extended empirical analysis on FluSight.

\begin{enumerate}[1.]

  \item The authors mention the difficulty of missing submissions during the FluSight season (hence the
    display chosen for Figure 10). How was this handled for Table 2? Do the averages refer to different
    sets of forecasting tasks for each model? If so the results are not fully comparable. The referenced
    paper by Cramer et al contains a (heuristic) scheme to compare scores in such a setting, which
    resembles what was done for Figure 10. This may be more appropriate than reporting raw averages.
    Models with very small submission fractions (maybe those with less than 80\%, definitely those with
    less than 50\%) in my opinion should just be removed altogether.
    \\~\\
    \emph{The columns RLWIS and RWIS in table 2 correspond directly to the 
          values plotted in figure 10. We rearranged the table so that RLWIS and
          RWIS come first. The updated table, shown as "Table 1" is below.
          To clarify what's given in table 2 we stated the
          following.}

    \begin{quote}
      Table 2 gives the numerical values for RLWIS 
      and relative WIS (RWIS) over
      all locations and weeks for each model, relative to the ASGD\_NORM2 model.
      It also includes the non relative mean WIS and LWIS for each model, that is the 
      overall mean scores for all forecasts submitted regardless of submission overlap
      with another model.
    \end{quote}

    \begin{quote}
      During the FluSight competition, most participating teams would occasionally 
      miss a forecast submission for some week, location, or horizon.
      The FluSight 
      models which are used for comparison are the
      13 non-ensemble models which submitted forecasts for at least 80\% of 
      FluSight targets during the 2023 season.

      Because forecast submissions for some targets were occassionally missed
      by participating teams, for any two models there was often a mismatch between
      targets for which forecasts were included or exlcuded. This makes a
      simple average of scores is an imperfect measure for comparison.
    \end{quote}

  \item It is uncommon to report median scores as their statistical properties are unclear. As mean score
    are also there I'm not fundamentally opposed to reporting them, but in my opinion mean scores
    alone are sufficient.  
    \\~\\
    \emph{Thank you for this comment. Because of it, we removed the median scores
    from the table which improved the aesthetic of the table as well as simplified
    it. See the new table below.}
\\~\\    
\end{enumerate}

\begin{table}[ht]
\centering
% \begin{adjustwidth}{-3cm}{-1.5cm}
\caption{Overall summary scores for each of the 13 non-ensemble FluSight
models with over 80\% submissions and 
the ASGD\_NORM2 and SIRD\_NORM2 models.
Summaries include 
mean weighted interval the relative log-weighted interval
and weighted interval scores (RLWIS/RWIS),
the overall mean log-weighted and weighted interval scores (MLWIS/MWIS), 
mean squared error
difference between predictive model empirical and theoretical coverage (MSEC)
see figure 9 c)), and percentage of forecasts 
submitted for all 
targets during the 2023 season (\% Forcs).}
\small
\begin{tabular}{lrrrrrrrr}
  % \hline
                              & RLWIS & RWIS & MLWIS & MWIS & MSEC  & \% Forcs\\ 
  \hline 
  ASGD\_NORM2                 & 1     & 1    & 0.34  & 62   & 0.001 & 100 \\ 
  PSI-PROF                    & 0.98  & 0.98 & 0.35  & 63   & 0.003 & 100 \\ 
  SIRD\_NORM2                 & 0.95  & 0.93 & 0.36  & 67   & 0.001 & 100 \\ 
  CEPH-Rtrend\_fluH           & 0.87  & 0.89 & 0.39  & 69   & 0.004 & 100 \\ 
  FluSight-baseline           & 0.83  & 0.73 & 0.41  & 85   & 0.023 & 100 \\ 
  UM-DeepOutbreak             & 0.74  & 0.73 & 0.46  & 83   & 0.010 & 99 \\ 
  cfarenewal-cfaepimlight     & 0.93  & 0.82 & 0.37  & 76   & 0.001 & 99 \\ 
  MOBS-GLEAM\_FLUH            & 0.97  & 0.95 & 0.35  & 66   & 0.021 & 98 \\ 
  LosAlamos\_NAU-CModel\_Flu  & 0.54  & 0.26 & 0.64  & 242  & 0.153 & 93 \\ 
  UGuelph-CompositeCurve      & 0.66  & 0.50 & 0.52  & 132  & 0.018 & 93 \\ 
  LUcompUncertLab-chimera     & 0.73  & 0.76 & 0.43  & 84   & 0.004 & 92 \\ 
  UGA\_flucast-INFLAenza      & 1.09  & 0.90 & 0.32  & 74   & 0.003 & 90 \\ 
  cfa-flumech                 & 0.86  & 0.65 & 0.41  & 101  & 0.019 & 89 \\ 
  UGA\_flucast-Copycat        & 0.99  & 0.89 & 0.35  & 76   & 0.012 & 87 \\ 
  GT-FluFNP                   & 0.74  & 0.56 & 0.48  & 105  & 0.046 & 86 \\ 
  % UNC\_IDD-InfluPaint         & 107 & 0.35 & 16 & 0.24 & 0.043 & 78 & 0.58 & 0.95 \\ 
  % NU\_UCSD-GLEAM\_AI\_FLUH    & 98 & 0.39 & 22 & 0.24 & 0.014 & 74 & 0.68 & 0.83 \\ 
  % SigSci-CREG                 & 59 & 0.49 & 12 & 0.31 & 0.018 & 71 & 0.62 & 0.74 \\ 
  % PSI-PROF\_beta              & 77 & 0.32 & 18 & 0.20 & 0.004 & 67 & 0.87 & 0.93 \\ 
  % NIH-Flu\_ARIMA              & 230 & 0.34 & 28 & 0.23 & 0.001 & 20 & 0.57 & 0.88 \\ 
  % GH-model                    & 161 & 1.52 & 28 & 1.47 & 0.373 & 17 & 0.18 & 0.27 \\ 
  % UGA\_flucast-OKeeffe        & 27 & 0.59 & 8 & 0.43 & 0.003 & 10 & 0.60 & 0.58 \\  
\end{tabular}
% \end{adjustwidth}
\label{tab:fin_analysis_stats}
\end{table}

\newpage

\Huge
\noindent R2
\normalsize
\\~\\
Forecasting Influenza Hospitaliztions Using a Bayesian Hierarchical Nonlinear 
Model with Discrepancy

This manuscript seems to be a very reasonable attempt at forecasting influenza 
induced hospitalizations. The
manuscript also puts the methodology in the context of the FluSight competition 
organised by the CDC in the US.
Apologies, I have found it hard to review as it is conceptually very hard to do 
so in my opinion. Firstly,
I do not find comfortable modelling the proportion, rather a natural outcome to 
model will be the number of hospitalization using a Generalized Linear model 
(Poisson distribution). Secondly, I find it un-intuitive to model the
flu curve using an Asymmetric Gaussian (ASG) Function. For a statistical model 
we focus on two aspects: the systematic part and the random part. Here, as far 
as  I understand the systematic part is the ASG function. The random
part has several components, which are hidden in Equation (5). My difficulty in 
modelling using the ASG
function lies in my inability to relate any meaningful meteorological variables 
(which are very much drivers of the seasonal flu outbreaks). See for example 
the paper,  "A hierarchical Bayesian model for improving short-term forecasting 
of hospital demand by including meteorological information" by  SK Sahu, 
B Baffour, PR Harper, JH Minty, C Sarran
Journal of the Royal Statistical Society, Series A (2014).

In spite of the above criticisms, I appreciate the hierarchical modeling 
approach of fitting the two models separately in components. But I wonder how 
the uncertainty has been propagated from one component model to another. 
Cutting the feedback idea is well understood but one still needs to integrate 
uncertainty to get an overall measure of the uncertainty in the forecasts. 
I did not clearly understand this issue from my reading of  the paper. I can 
suggest one way to propagate this uncertainty is by using MCMC samples 
(and methods), see for example, the paper,  "A rigorous statistical framework 
for spatio-temporal pollution prediction and estimation of its long-term impact 
on health" by 
D Lee, S Mukhopadhyay, A Rushworth, and SK Sahu, in Biostatistics (2017).

In conclusion, I do think that the authors have done a good job of analysis and 
simulation based demonstrations of the results. These can be further improved 
by including some better measures of assessing prediction/forecasting accuracy, 
e.g. the CRPS (Continuous Ranked Probability Score), see for example Chapter 6 
of the book
"Bayesian modeling of spatio-temporal data with R" by Sahu 
(Chapman and Hall, 2022). 
\\~\\
\emph{In the simulation study in section 4 of the first version of this paper, 
scoring was
done using the CRPS, but in the FluSight analysis in section 5 scoring was done
using the WIS because the available forecasts are not amenable to being scored
the with the CRPS. A recommendation from one of the reviewers was to simplify 
by choosing one or the other scores, and we chose the WIS. To respond to this 
point, however, we modified the first paragraph on page 17. The paragraph is 
below and the added statements are bolded.
}

\begin{quote}
Model comparison was done by calculating a proper scoring rule for each 
forecast.
Proper scoring rules are the 
current standard for comparing performance between probabilistic forecasts 
and selecting the best forecasts according to the notion of maximizing 
sharpness subject to (auto-)calibration 
\cite[]{gneiting2007probabilistic, tsyplakov2013evaluation}. Proper scoring 
rules are commonly used in forecast comparison, and they are designed such 
that a 
forecaster is incentivized to be honest in the reporting of their forecasts 
\cite[]{gneiting2007strictly, gneiting2014probabilistic}.
\textbf{Commonly used proper scoring rules for continuous distribution forecasts
include the log score and the continuously
ranked probability score (CRPS).} 
The proper scoring rule primarily used in the FluSight competition and the 
COVID-19 Forecast Hub, and which is used in this simulation study,
is the negatively oriented
(smaller is better)
weighted interval score (WIS)
\cite[]{mathis2024evaluation, bracher2021evaluating}.
The WIS is used for scoring quantile or interval 
forecasts \textbf{--the forecast format used in both FluSight and the 
COVID-19 Forecast Hub--}
or forecasts made up of predicitive intervals of several nominal 
levels
\cite[]{gneiting2007strictly, gneiting2014probabilistic, bracher2021evaluating}. 
The WIS is 
defined in (\ref{eq:wis2}) where $Q$ is a forecast represented by all 
included quantiles, $B$ is the number of intervals,  $y^*$ is the observed 
value targeted by the forecast, $w_0 = 1/2$ and $w_b = \alpha_b / 2$ are 
weights for each interval, and $\alpha_b$ is the nominal level of the $b^{th}$ 
interval where $b \in \{1, ..., B\}$. 
$IS_{\alpha}$ is the interval score (IS), a proper scoring rule for a 
single interval, defined in (\ref{eq:is}). Here 
$\vmathbb{1}\{\cdot\}$ is the indicator function.
\textbf{We note that as the number of forecast intervals $B$ increases, the WIS becomes 
arbitrarily close to
the CRPS. Thus, forecast scores under the CRPS and the WIS should be similar
\cite[]{bracher2021evaluating}.}
\end{quote}

\bibliographystyle{ba}
\bibliography{master_bib}

\end{document}